\chapter{Installation}
\label{chInstallation}

This section describes the installation procedure for the RoboCup 3D Simspark simulation server and for the \libbats library.

\section{RoboCup 3D Simspark Simulation Server}

For the latest installation instructions of the simulator, see the SimSpark project's wiki: \\
\url{http://simspark.sourceforge.net/wiki}\\
There you will find instructions for installation on various Linux distributions, as well as for Windows and MacOS.

\section{libbats}

\begin{description}
  \item[Dependencies] Install the necessary dependencies. For Ubuntu, and possibly other Debian based distributions, run:
\begin{verbatim}
$ sudo apt-get install libxml2-dev libsigc++-2.0-dev libgtkmm-2.4-dev
\end{verbatim}
On Arch, run as root:
\begin{verbatim}
$ pacman -S libxml2 libsigc++ gtkmm
\end{verbatim}
  \item[libbats] Install the latest version of {\tt libbats}:
    \begin{enumerate}
      \item Download the libbats-x.y.z.tar.gz (where x.y.z is the latest version number, currently 2.1.0) source code package from: \\
        \url{http://launchpad.net/littlegreenbats/+download}
      \item Unpack and navigate to the source code directory:
\begin{verbatim}
$ tar xvzf libbats-x.y.ztar.gz
$ cd libbats-x.y.z
\end{verbatim}
      \item Configure, make and install:
\begin{verbatim}
$ ./configure
$ make
$ sudo make install
\end{verbatim}
    \end{enumerate}
\end{description}

\section{RoboViz}

RCSSServer3D comes with a monitor, RCSSMonitor3D, however it is rather basic, both graphically as functionally. We suggest using RoboViz instead, which offers much better visualization, and an advanced debugging interface that is supported by \libbats. Installation instructions for RoboViz can be found at:\\
\url{https://sites.google.com/site/umroboviz/}

\section{Testing}

To test whether everything is working correctly, start the simulator with:
\begin{verbatim}
$ rcssserver3d
\end{verbatim}
This command should give you a greeting saying something similar to the following:
\begin{verbatim}
rcssserver3d (formerly simspark), a monolithic simulator 0.6.5
Copyright (C) 2004 Markus Rollmann, 
Universität Koblenz.
Copyright (C) 2004-2009, The RoboCup Soccer Server Maintenance Group.

Type '--help' for further information
\end{verbatim}
plus some initialization output. This output will contain some messages such as `{\tt ERROR: cannot find TextureServer}' and `{\tt ERROR: no FPSController found at}\\{\tt '/usr/scene/camera/physics/controller'}'; you can ignore these messages.

Next start RoboViz. Change to the directory containing RoboViz' binary and run:
\begin{verbatim}
./roboviz.sh
\end{verbatim}
After some time in which RoboViz connects to the server and initializes its graphics a green field will appear with some lines and two goals. Ta-da, you are successfully running the simulator! If RoboViz reports 'Disconnected', you are not and something went wrong with installing the simulator; see below for some troubleshooting tips.

Finally start the example agent supplied with \libbats; in the \libbats source directory execute:
\begin{verbatim}
$ cd examples/helloworld
$ ./helloworld
\end{verbatim}
If everything went well, an agent should appear, standing in the left side of the field, waving its arms.

\section{Troubleshooting}
\begin{enumerate}
    \item To uninstall Simspark or RCSSServer3D, simply execute the following command within the respective build directory:
\begin{verbatim}
$ sudo make uninstall
\end{verbatim}
    \item If compilation fails due to missing dependencies, as reported when running {\tt cmake} for the simulator or {\tt configure} for \libbats, try looking up the required libraries using your distributions's package manager. For Ubuntu:
\begin{verbatim}
$ sudo apt-cache search <missing package>
\end{verbatim}
For Arch:
\begin{verbatim}
$ pacman -Ss <missing package>
\end{verbatim}
    \item Please refer to the SimSpark wiki for more information on how to install (or use) simspark, e.g. when using a different operating system: \\
        \url{http://simspark.sourceforge.net/wiki/index.php/Main_Page}.
    \item Search through the sserver-three-d mailing list to see if there is a solution to your problem:\\
    \url{http://sourceforge.net/search/?type_of_search=mlists&group_id=24184}\\
    otherwise post it to the list (see \url{http://sourceforge.net/mail/?group_id=24184}).
    \item For problems with \libbats, see the Bugs and Answers sections at: \\
    \url{http://launchpad.net/littlegreenbats}.
\end{enumerate}

